---
title: "COVID-19 report for `r params$country` in `r params$year`"
output:
  html_document:
    theme: journal
    highlight: tango
params:
  country: "Netherlands"
  year: 2022
  months: [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(here)
library(conflicted)
conflicts_prefer(dplyr::filter)
```


## Data Overview

We are generating a COVID-19 report for `r params$country` for the year `r params$year` and for the months `r paste(params$months, collapse = ', ')`.

```{r Importing data, echo=FALSE, eval=TRUE}
df <- read.csv("https://opendata.ecdc.europa.eu/covid19/nationalcasedeath_eueea_daily_ei/csv", na.strings = "", fileEncoding="UTF-8-BOM")

# Process data
covid_data <- df %>% filter(countriesAndTerritories == params$country,
                            year == params$year,
                            month %in% params$months)
```


```{r stats, echo=FALSE, eval=TRUE}
# Summary statistics of the COVID-19 cases
summary_cases <- covid_data %>%
  summarize(total_cases = sum(cases, na.rm = TRUE %>% round(digits = 2)),
            average_cases = mean(cases, na.rm = TRUE %>% round(digits = 2)),
            max_cases = max(cases, na.rm = TRUE %>% round(digits = 2)),
            min_cases = min(cases, na.rm = TRUE %>% round(digits = 2)))

summary_deaths <- covid_data %>%
  summarize(total_deaths = sum(deaths, na.rm = TRUE %>% round(digits = 2)),
            average_deaths = mean(deaths, na.rm = TRUE %>% round(digits = 2)),
            max_deaths = max(deaths, na.rm = TRUE%>% round(digits = 2)),
            min_deaths = min(deaths, na.rm = TRUE%>% round(digits = 2)))
```

## COVID-19 Cases 

```{r cases, echo=FALSE}
covid_data %>% ggplot(aes(x = month, y = log10(cases + 1))) +
  geom_col(na.rm = T) +
  scale_x_continuous(breaks = seq(1, max(12), by = 1)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
    labs(title = paste("COVID-19 Cases in", params$country, "in", params$year),
       x = "Months",
       y = "Number of Cases (Log 10 Scale)")
```

### Summary Statistics

-   **Total COVID-19 cases = `r summary_cases$total_cases`**

-   **Average COVID-19 cases per day = `r summary_cases$average_cases`**


## COVID-19 Deaths

```{r Deaths, echo=FALSE}
covid_data %>% ggplot(aes(x = month, deaths)) +
  geom_col(na.rm = T) +
  scale_x_continuous(breaks = seq(1, max(12), by = 1)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
    labs(title = paste("COVID-19 Deaths in", params$country, "in", params$year),
       x = "Months",
       y = "Number of Deaths")
```

### Summary Statistics

-   **Total COVID-19 deaths = `r summary_deaths$total_deaths`**

-   **Average COVID-19 deaths per day = `r summary_deaths$average_deaths`**


